A comparison of different methods for calculating tangent-stiffness matrices in a massively parallel computational peridynamics code.

### 2013-09-25 jacobian_calc_CMAME.tex
added reference for LPS model

After the bulk of the study had been conducted, gained familiarity with \emph{Peridigm} revealed
that the implementations of forward-difference, central-difference and by virtue of reusing code,
complex-step, in \emph{Peridigm} were not optimal due to inefficiencies in the implementations of
the Linear Peridynamic Solid (LPS) model. Each of these implementations included a large amount of
unnecessary re-calculation of constant values.  The approach used does minimize temporary memory
use, and allow other objects in \emph{Peridigm} to make calls directly to methods in the material
objects, but is wastefull. Unpublished preliminary results show that when calculations for the
constant quantity 'zeta' in the LPS implementation are re-used and only perturbed and previously
pertubed functions of independent variables are recomputed throughout the course of a Jacobian
evaluation, complex-step performs at the speed of unmodified forward-difference. Because of this it
is speculated that forward-difference and central-difference could be modified in nearly the same
way to perform more quickly than AD. In fact, because central-difference involves double the number
of model evaluations than that of complex-step, central-difference stands to gain proportionally
more efficiency. Although no formal academic paper will be likely written based on this, a brief
report and an enhanced performance implementation of finite-difference for peridynamic LPS Jacbian
calculation  in \emph{Peridigm} will eventually be made available on the corresponding author's
website and is intended to be submitted for official inclusion in future releases of \emph{Peridigm}.

>>>>

After the bulk of the study had been conducted, gained familiarity with \emph{Peridigm} revealed
that the implementations of forward-difference, central-difference and by virtue of reusing code,
complex-step, in \emph{Peridigm} were not optimal due to inefficiencies in the implementations of
the Linear Peridynamic Solid (LPS) model \cite[A 1]{ref-LPS}. Each of these implementations included a large amount of
unnecessary re-calculation of constant values.  The approach used does minimize temporary memory
use, and allow other objects in \emph{Peridigm} to make calls directly to methods in the material
objects, but is wastefull. Unpublished preliminary results show that when calculations for the
constant quantity 'zeta' in the LPS implementation are re-used and only perturbed and previously
pertubed functions of independent variables are recomputed throughout the course of a Jacobian
evaluation, complex-step performs at the speed of unmodified forward-difference. Because of this it
is speculated that forward-difference and central-difference could be modified in nearly the same
way to perform more quickly than AD. In fact, because central-difference involves double the number
of model evaluations than that of complex-step, central-difference stands to gain proportionally
more efficiency. Although no formal academic paper will be likely written based on this, a brief
report and an enhanced performance implementation of finite-difference for peridynamic LPS Jacbian
calculation  in \emph{Peridigm} will eventually be made available on the corresponding author's
website and is intended to be submitted for official inclusion in future releases of \emph{Peridigm}.
